### 2016-12-20 paper/chapters/gan.tex
more stuff on GANS

%%% when introduced, game theory approach
Generative Adversarial Networks (GAN) is a framework recently proposed by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
%which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).

The idea is to have two players playing a non-cooperative game against each other.
One player is called the generator $G$ tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually originating from a dataset.
It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.

Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
On the other side, $G$ tries to produce as many samples as possibly which got accepted as real by $D$.\\

@@

To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.

In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
GANs uses neural network as function approximator for $G$ as well as $D$.
The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
where $\xi$ denotes the parameters learned by the neural network.\\
Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.

@@

\begin{figure}[htb]
\centering
\includestandalone[width=\linewidth,mode=buildnew]{media/gan_conceptual}
  \caption{Conceptual figure of generative adversarial network}
  \label{fig:gan_a}
  \medskip
  \small
  The above figure is illustrating the information and gradient flow in the GAN framework.\\
  \emph{data space} denotes the real-world data sampled from $p_{data}$, which we aim to model.
  Dashed arrows denote the gradient flow.
\end{figure}

@@

\subsection{Architecture}
\label{sub:gan_arch}
Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).

@@

%Another common form of the generator loss is $\mathcal{L'}_g$.
%$$
$$
\mathcal{L}_g = \min \log(1 - D(G(z)))\\
\mathcal{L'}_g = \max \log D(G(z))
$$
This formulation has the benefit of the same optimum fixed point, but has a better gradient flow early in training when $G$ produces samples which can be distinguished easily from real data by $D$.

@@

\subsection{Training}
\label{sub:gan_training}
Combining the loss functions of both generator and discriminator yields us the following objective with value function $V(G,D)$.
$$
\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
$$

>>>>

%%% when introduced, game theory approach
Generative Adversarial Networks (GAN) is a framework proposed in 2014 by Goodfellow et al. \cite{gan:2014} based on a game theoretic approach to generative modelling.
%which has gathered a lot of attention in the deep learning community and in the last few years many extensions have been [proposed] (citations).

GANs circumvent the issues arising the inference problem by playing a two-player game. Both players play non-cooperatively against each other, which means they only try to maximize their respective success or profit.
One player is called the generator $G$ which tries to produce authentic looking data samples while the other player, the discrimininator $D$ has to decide whether a received sample is generated or actually real data.
It is important to note that the generator gets to know whether the discriminator accepted the generated output as \emph{real} or rejected it as \emph{fake}.

Each player plays to maximize its own success, which means the discriminator maximizes the number of samples which got discriminated correctly.
On the other side, $G$ tries to produce as many real-looking samples as possible to get accepted by $D$.\\

@@

Another formulation of the GAN framework is to define the zero-sum game between the generative model and its adversary.
The payoff of the discriminator can be described as the value function $v(G,D)$ while the payoff of the generator is the negative payoff of $D$: $-v(G,D)$.\\
This formulation allows us to define the convergence as equation \ref{zerosum_convergence}.
\begin{equation}
  \label{eq:zerosum_convergence}
  g^* = \arg \min_{G} \max_{D} v(g,d)
\end{equation}

@@

\begin{figure}[htb]
\centering
\includestandalone[width=\textwidth,mode=buildnew]{media/gan_conceptual}
  \caption{Architecture used for generative adversarial networks}
  \label{fig:gan_a}
  \medskip
  \small
  The above figure illustrates the computational graph and gradient flow in the GAN framework.\\
  $X$ denotes the dataset containing real-world samples from $p_{data}$, which we aim to model.\\
  Solid arrows indicate the forward pass in the model while dashed arrows denote the gradient flow during a backward pass.
\end{figure}

@@

\subsection{Architecture}
\label{sub:gan_arch}
To formalize the setting in which the game between $G$ and $D$ takes place, we define the true, however unknown data distribution $p_{data}(x)$ from which samples are available as dataset $X=\{x^{(1)}, x^{(2)}, \dots, x^{(N)}\}$.

In the case of generative modelling, the goal is to approximate $p_{data}(x)$ with $p_{model}(x)$.
GANs uses neural network as function approximator for $G$ as well as $D$.
The generator network $G$ produces samples from $p_{model}(x)$ by transforming samples from random noise as $x = G_\xi(z)$,
where $\xi$ denotes the parameters learned by the neural network representing $G$.\\
Competing with $G$ is the discriminator $D$ trained to distinguish samples $ \tilde{x} \sim p_{model}(x)$ from real data $x \sim p_{data}(x)$.\\

Figure \ref{fig:gan_a} summarizes the described architecture.\\
%Formalizing the setup in a probabilistic way, we have the unknown data distribution $p_{data}(x)$ from which the training examples have been drawn and the marginalized model distribution $p_{model}(x) = \int p(z)p(x|z) \textit{d}z$ over the noise space $z$ (also latent space).

@@

%Another common form of the generator loss is $\mathcal{L'}_g$.
%$$
\begin{equation}
  \label{eq:l_g_1}
  \mathcal{L}_g = \min \log(1 - D(G(z)))
\end{equation}

However, for gradient-based learning $\mathcal{L}_g$ provides poor gradient during learning where the discriminator is able to reject samples from the generator with high confidence.\\
Therefore is in practice often equation \ref{eq:l_g_2} instead of equation \ref{eq:l_g_1} to optimize its objective.\\
\begin{equation}
  \label{eq:l_g_2}
  \mathcal{L'}_g = \max \log D(G(z))
\end{equation}

@@

\subsection{Training}
\label{sub:gan_training}
\begin{figure}[htb]
\centering
  \includegraphics[width=\linewidth]{media/gan_training-crop}
  \caption{Architecture used for generative adversarial networks}
  \label{fig:gan_training}
  \medskip
  \small
\end{figure}

Combining the objective functions of both generator and discriminator in equation \ref{eq:l_g_2} and \ref{eq:l_d}, respectively yields us the following objective with value function $V(G,D)$.
$$
\min_G \max_D V(D,G) = \mathbb{E}_{x \sim p_{data}}[\log D(x)] + \mathbb{E}_{z \sim p_z(z)}[\log(1 - D(G(z)))]
$$
